Synthesizing Symmetric Lenses

### 2018-06-30 papers/SymmetricLenses/main.tex
fixed typo
in progress edits to section 2

In this section, we give an overview of simple symmetric lenses and
our corresponding synthesis algorithm using an example drawn from a
hypothetical U.S. company.  In this company, human resources and upper
management store information about employees in separate files:
Management records the names of employees and their salaries while
Human Resources (HR) stores the names of employees and their health
insurance providers.  Figure~\ref{fig:minimized-representations} show
two example files and regular expressions describing the formats.
The company wants to keep these two files synchronized:
if management adds a new employee, ``Chris Roe: 32500'', 
the corresponding entry ``Chris Roe, UNK'' should be added to HR's file.

Furthermore, if HR
changes ``John Public'' to ``Jon Public'', that name change should be reflected
in both files. However, not all the data should be synchronized. The HR file
should not get altered due to salary changes, and the management file should not
get altered due to insurance company changes. While names should be matched
across the two formats, insurance company and salary information should be left
unmatched. Such a synchronization can be achieved with simple symmetric lenses.

@@

Lenses synchronize files by providing 4 functions, two \emph{create} functions,
and two \emph{put} functions. Create functions are used to initialize data in
one format when provided with data in the other. If a company.mgmt file is
initialized, a synchronized company.hr file should also be created, and vice
versa. Put functions are used to propagate edits from one format to the other.
If a company.hr file is updated, we should be able to combine it with a
non-updated company.mgmt file, and output a new company.mgmt file, synchronized
to the updated company.hr. To have a simple symmetric lens is to have these
required synchronization functions. Formally, a simple symmetric lens between
$X$ and $Y$ comprises 4 functions:
\begin{enumerate}
\item $\CreateR \OfType X \rightarrow Y$
\item $\CreateL \OfType Y \rightarrow X$
\item $\PutR \OfType X \rightarrow Y \rightarrow Y$
\item $\PutL \OfType Y \rightarrow X \rightarrow X$
\end{enumerate}
Furthermore, if a file is not edited, but is then resynchronized using a put
function, no data should be changed. To be a simple symmetric lens, one must
satisfy the following ``round-tripping'' laws, guaranteeing identity edits won't
change the file data.
\begin{gather}
  \tag{\CreatePutRL}
  \PutLOf{(\CreateROf{x})}{x} = x\\
  \tag{\CreatePutLR}
  \PutROf{(\CreateLOf{y})}{y} = y\\
  \tag{\PutRL}
  \PutLOf{(\PutROf{x}{y})}{x} = x\\
  \tag{\PutLR}
  \PutROf{(\PutLOf{y}{x})}{y} = y
\end{gather}

@@

Consider the trivial lens
\lstinline{disc(emp_salary,emp_row,rep(emp_salary),rep(emp_row))}. As this is
the disconnect lens, the cost of each side is the entropy of the data format, so
the total cost is $\EntropyOf{\lstinline{emp_salary}} +
\EntropyOf{\lstinline{emp_row}}}$. \lstinline{emp_salary} contains strictly more
imformation than \EntropyOf{\lstinline{": " . salary}}, and \lstinline{emp_row}
contains strictly more information than \EntropyOf{\lstinline{": " . salary}},
so by our metric \lstinline{single_emp_map} is preferable to the disconnect
lens.
% end overview

>>>>

In this section, we give an overview of simple symmetric lenses and
our corresponding synthesis algorithm using an example drawn from a
hypothetical U.S. company.  In this company, management and
human resources (HR) store information about employees in separate files:
management records the names of employees and their salaries while
HR stores the names of employees and their health
insurance providers.  Figure~\ref{fig:minimized-representations} gives
examples of the two types of files and regular expressions describing
the two formats.
The company uses simple symmetric lenses to keep these files synchronized.
If management adds a new employee, ``Chris Roe: 32500'',  a lens adds 
the corresponding entry ``Chris Roe, UNK'' to HR's file.  The
sentinel value ``UNK'' represents the fact that the employee's
insurance company is currently unknown.  A similar update happens if
HR adds a new employee before management, in which case the sentinel
value ``N/A'' reflects the unknown salary information. 
Furthermore, if HR
corrects an error in an employee's name, say changing ``John Public''
to ``Jon Public'', the lens mirrors that name change in management's
file. Not all the data is mirrored, however. The management file 
is not updated in response to insurance changes, and the HR file
is oblivious to salary changes.  Simple symmetric lenses are
appropriate for keeping these two files synchronized because they have
a mix of shared and unshared information.

@@

Simple symmetric lenses synchronize files using four functions,
two \emph{create} functions, 
and two \emph{put} functions. Create functions are used to initialize data in
one format when provided with data in the other. If a company.mgmt file is
initialized, a synchronized company.hr file should also be created, and vice
versa. Put functions are used to propagate edits from one format to the other.
If a company.hr file is updated, we should be able to combine it with a
non-updated company.mgmt file, and output a new company.mgmt file, synchronized
to the updated company.hr. To have a simple symmetric lens is to have these
required synchronization functions. Formally, a simple symmetric lens between
$X$ and $Y$ comprises 4 functions:
\begin{enumerate}
\item $\CreateR \OfType X \rightarrow Y$
\item $\CreateL \OfType Y \rightarrow X$
\item $\PutR \OfType X \rightarrow Y \rightarrow Y$
\item $\PutL \OfType Y \rightarrow X \rightarrow X$
\end{enumerate}
Furthermore, if a file is not edited, but is then resynchronized using a put
function, no data should be changed. To be a simple symmetric lens, one must
satisfy the following ``round-tripping'' laws, guaranteeing identity edits won't
change the file data.
\begin{gather}
  \tag{\CreatePutRL}
  \PutLOf{(\CreateROf{x})}{x} = x\\
  \tag{\CreatePutLR}
  \PutROf{(\CreateLOf{y})}{y} = y\\
  \tag{\PutRL}
  \PutLOf{(\PutROf{x}{y})}{x} = x\\
  \tag{\PutLR}
  \PutROf{(\PutLOf{y}{x})}{y} = y
\end{gather}

@@

Consider the trivial lens
\lstinline{disc(emp_salary,emp_row,rep(emp_salary),rep(emp_row))}. As this is
the disconnect lens, the cost of each side is the entropy of the data format, so
the total cost is $\EntropyOf{\lstinline{emp_salary}} +
\EntropyOf{\lstinline{emp_row}}$. \lstinline{emp_salary} contains strictly more
imformation than \EntropyOf{\lstinline{": " . salary}}, and \lstinline{emp_row}
contains strictly more information than \EntropyOf{\lstinline{": " . salary}},
so by our metric \lstinline{single_emp_map} is preferable to the disconnect
lens.
% end overview
